### 2018-10-01 post-linearity-2018/sequent.tex
Proof for cut-elimination in the appendix

The syntax for $\cat{L}$-types are of the typical form for the Lambek
Calculus. We have a unit type $[[UnitS]]$, a non-commutative tensor product
$[[A (>) B]]$, right implication $[[A -> B]]$, and left implication
$[[B <- A]]$.

The sequenct calculus for CNC logic can be found in
Figure~\ref{fig:CNC-sequent-calculus}.
\begin{figure}[!h]
  \footnotesize
  \begin{tabular}{|c|}
    \hline\\
    \begin{mathpar}
    \SCdruleTXXax{} \and
    \SCdruleTXXunitL{} \and
    \SCdruleTXXunitR{} \and
    \SCdruleTXXtenL{} \and
    \SCdruleTXXtenR{} \and
    \SCdruleTXXimpL{} \and
    \SCdruleTXXimpR{} \and
    \SCdruleTXXGr{} \and
    \SCdruleTXXex{} \and
    \SCdruleTXXcut{}
    \end{mathpar}\\\\
    \hline
    \\[5px]
    \begin{mathpar}
    \SCdruleSXXax{} \and
    \SCdruleSXXunitLOne{} \and
    \SCdruleSXXunitLTwo{} \and
    \SCdruleSXXunitR{} \and
    \SCdruleSXXex{} \and
    \SCdruleSXXtenLOne{} \and
    \SCdruleSXXtenLTwo{} \and
    \SCdruleSXXtenR{} \and
    \SCdruleSXXimpL{} \and
    \SCdruleSXXimprL{} \and
    \SCdruleSXXimprR{} \and
    \SCdruleSXXimplL{} \and
    \SCdruleSXXimplR{} \and
    \SCdruleSXXFl{} \and
    \SCdruleSXXFr{} \and
    \SCdruleSXXGl{} \and
    \SCdruleSXXcutOne{} \and
    \SCdruleSXXcutTwo{} \and
    \end{mathpar}\\\\
    \hline
  \end{tabular}
  \caption{Sequent Calculus for CNC Logic}
  \label{fig:CNC-sequent-calculus}
\end{figure}
We split the figure in two: the top of the figure are the rules of
intuitionistic linear logic whose sequents are the $\mathcal{C}$-sequents
denoted by $[[P |-c X]]$, and the bottom of the figure are the rules for the
mixed commutative/non-commutative Lambek calculus whose sequents are the
$\mathcal{L}$-sequents denoted by $[[G |-l A]]$, but the two halves are
connected via the functor rules $\SCdruleTXXGrName{}$,
$\SCdruleSXXGlName{}$, $\SCdruleSXXFlName{}$, and $\SCdruleSXXFrName{}$, and
the rules $\SCdruleSXXunitLOneName{}$, $\SCdruleSXXexName{}$,
$\SCdruleSXXtenLOneName{}$, $\SCdruleSXXimpLName{}$,
$\NDdruleSXXcutOneName{}$.

We prove cut elimination for the sequent calculus. We define the
\textit{degree} $|X|$ (resp. $|A|$) of a commutative (resp. non-commutative)
formula to be the number of logical connectives in the proposition. For
instance, $|[[X (*) Y]]| = |[[X]]| + |[[Y]]| + 1$. The \textit{cut rank}
$c(\Pi)$ of a proof $\Pi$ is one more than the maximum of the ranks of all
the cut formulae in $\Pi$, and $0$ if $\Pi$ is cut-free. Then the
\textit{depth} $d(\Pi)$ of a proof $\Pi$ is the length of the longest path
in the proof tree (so the depth of an axiom is $0$). The key to the proof
of cut elimination is the following lemma, which shows how to transform a
single cut, either by removing it or by replacing it with one or more
simpler cuts.

\begin{lemma}[Cut Reduction]
  \label{lem:cut-reduction}
  \begin{enumerate}
  \item If $\Pi_1$ is a proof of $[[I |-c X]]$ and $\Pi_2$ is a proof of
  $[[P1,X,P2 |-c Y]]$ with $c(\Pi_1)$, $c(\Pi_2)\leq |X|$, then there exists
  a proof $\Pi$ of $[[P1, I, P2 |-c Y]]$ with $c(\Pi)\leq |X|$.
  \item If $\Pi_1$ is a proof of $[[I |-c X]]$ and $\Pi_2$ is a proof of
  $[[G1; X; G2 |-l A]]$ with $c(\Pi_1)$, $c(\Pi_2)\leq |X|$, then there
  exists a proof $\Pi$ of $[[G1; I; G2 |-l A]]$ with $c(\Pi)\leq |X|$.
  \item If $\Pi_1$ is a proof of $[[G |-l A]]$ and $\Pi_2$ is a proof of
  $[[D1; A; D2 |-l B]]$ with $c(\Pi_1)$, $c(\Pi_2)\leq |A|$, then there
  exists a proof $\Pi$ of $[[D1; G; D2 |-l B]]$ with $c(\Pi)\leq |A|$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  The proof is done case by case similar as in \cite{Mellies:2009}.
\end{proof}

@@

\begin{theorem}[Cut Elimination]
  Let $\Pi$ be a proof of a sequent $[[I |-c X]]$ or $[[G |-l A]]$ s.t.
  $c(\Pi)>0$. Then there is an algorithm which yields a cut-free proof
  $\Pi'$ of the same sequent.
\end{theorem}
\begin{proof}
  This follows immediately by induction on $c(\Pi)$ and
  Lemma~\ref{lem:less-cut-rank}.
\end{proof}

>>>>

The syntax for $\cat{L}$-types are of the typical form for the Lambek
Calculus. We have two unit types $[[UnitS]]$ (one for each side), a
non-commutative tensor product $[[A (>) B]]$, right implication
$[[A -> B]]$, and left implication $[[B <- A]]$.

The sequenct calculus for CNC logic can be found in
Figure~\ref{fig:CNC-sequent-calculus}. We split the figure in two: the top
of the figure are the rules of intuitionistic linear logic whose sequents
are the $\mathcal{C}$-sequents denoted by $[[P |-c X]]$, and the bottom of
the figure are the rules for the mixed commutative/non-commutative Lambek
calculus whose sequents are the $\mathcal{L}$-sequents denoted by
$[[G |-l A]]$, but the two halves are connected via the functor rules
$\SCdruleTXXGrName{}$, $\SCdruleSXXGlName{}$, $\SCdruleSXXFlName{}$, and
$\SCdruleSXXFrName{}$, and the rules $\SCdruleSXXunitLOneName{}$,
$\SCdruleSXXexName{}$, $\SCdruleSXXtenLOneName{}$, $\SCdruleSXXimpLName{}$,
$\NDdruleSXXcutOneName{}$.
\begin{figure}[!h]
  \footnotesize
  \begin{tabular}{|c|}
    \hline\\
    \begin{mathpar}
    \SCdruleTXXax{} \and
    \SCdruleTXXunitL{} \and
    \SCdruleTXXunitR{} \and
    \SCdruleTXXtenL{} \and
    \SCdruleTXXtenR{} \and
    \SCdruleTXXimpL{} \and
    \SCdruleTXXimpR{} \and
    \SCdruleTXXGr{} \and
    \SCdruleTXXex{} \and
    \SCdruleTXXcut{}
    \end{mathpar}\\\\
    \hline
    \\[5px]
    \begin{mathpar}
    \SCdruleSXXax{} \and
    \SCdruleSXXunitLOne{} \and
    \SCdruleSXXunitLTwo{} \and
    \SCdruleSXXunitR{} \and
    \SCdruleSXXex{} \and
    \SCdruleSXXtenLOne{} \and
    \SCdruleSXXtenLTwo{} \and
    \SCdruleSXXtenR{} \and
    \SCdruleSXXimpL{} \and
    \SCdruleSXXimprL{} \and
    \SCdruleSXXimprR{} \and
    \SCdruleSXXimplL{} \and
    \SCdruleSXXimplR{} \and
    \SCdruleSXXFl{} \and
    \SCdruleSXXFr{} \and
    \SCdruleSXXGl{} \and
    \SCdruleSXXcutOne{} \and
    \SCdruleSXXcutTwo{} \and
    \end{mathpar}\\\\
    \hline
  \end{tabular}
  \caption{Sequent Calculus for CNC Logic}
  \label{fig:CNC-sequent-calculus}
\end{figure}

We prove cut elimination for the sequent calculus. We define the
\textit{degree} $|X|$ (resp. $|A|$) of a commutative (resp. non-commutative)
formula to be the number of logical connectives in the proposition. For
instance, $|[[X (*) Y]]| = |[[X]]| + |[[Y]]| + 1$. The \textit{cut rank}
$c(\Pi)$ of a proof $\Pi$ is one more than the maximum of the ranks of all
the cut formulae in $\Pi$, and $0$ if $\Pi$ is cut-free. Then the
\textit{depth} $d(\Pi)$ of a proof $\Pi$ is the length of the longest path
in the proof tree (so the depth of an axiom is $0$). The key to the proof
of cut elimination is the following lemma, which shows how to transform a
single cut, either by removing it or by replacing it with one or more
simpler cuts.
\begin{lemma}[Cut Reduction]
  \label{lem:cut-reduction}
  \begin{enumerate}
  \item If $\Pi_1$ is a proof of $[[I |-c X]]$ and $\Pi_2$ is a proof of
  $[[P1,X,P2 |-c Y]]$ with $c(\Pi_1)$, $c(\Pi_2)\leq |X|$, then there exists
  a proof $\Pi$ of $[[P1, I, P2 |-c Y]]$ with $c(\Pi)\leq |X|$.
  \item If $\Pi_1$ is a proof of $[[I |-c X]]$ and $\Pi_2$ is a proof of
  $[[G1; X; G2 |-l A]]$ with $c(\Pi_1)$, $c(\Pi_2)\leq |X|$, then there
  exists a proof $\Pi$ of $[[G1; I; G2 |-l A]]$ with $c(\Pi)\leq |X|$.
  \item If $\Pi_1$ is a proof of $[[G |-l A]]$ and $\Pi_2$ is a proof of
  $[[D1; A; D2 |-l B]]$ with $c(\Pi_1)$, $c(\Pi_2)\leq |A|$, then there
  exists a proof $\Pi$ of $[[D1; G; D2 |-l B]]$ with $c(\Pi)\leq |A|$.
  \end{enumerate}
\end{lemma}
The proof for Lemma~\ref{lem:cut-reduction} is done case by case on the last
step of $\Pi_1$ and $\Pi_2$ and by induction on $d(\Pi_1)$ and $d(\Pi_2)$,
following \cite{Mellies:2009}. For instance, suppose $\Pi_1$ is a proof of
$[[I1, X2, X1, I2 |-c Y]]$ and $\Pi_2$ is a proof of $[[P1, Y, P2 |-c Z]]$. 
Consider the case where the last step in $\Pi_1$ uses the rule
$\NDdruleTXXbetaName{}$. $\Pi_1$ can be depicted as follows, where the
previous steps is denoted by $\pi$:
\begin{center}
  \scriptsize
  $\Pi_1$:
  \begin{math}
    $$\mprset{flushleft}
    \inferrule* [right={\tiny ex}] {
      {
        \begin{array}{c}
          \pi \\
          {[[I1, X1, X2, I2 |-c Y]]}
        \end{array}
      }
    }{[[I1, X2, X1, I2 |-c Y]]}
  \end{math}
\end{center}
By assumption, $c(\Pi_1),c(\Pi_2)\leq |Y|$. By induction on $\pi$ and
$\Pi_2$, there is a proof $\Pi'$ for the sequent \\
$[[P1, I1, X1, X2, I2, P2 |-c Z]]$ s.t. $c(\Pi')\leq|Y|$. Therefore, the
proof $\Pi$ can be constructed as follows, and $c(\Pi)=c(\Pi')\leq|Y|$.
\begin{center}
  \scriptsize
  \begin{math}
    $$\mprset{flushleft}
    \inferrule* [right={\tiny ex}] {
      {
        \begin{array}{c}
          \Pi' \\
          {[[P1, I1, X1, X2, I2, P2 |-c Z]]}
        \end{array}
      }
    }{[[P1, I1, X2, X1, I2, P2 |-c Z]]}
  \end{math}
\end{center}
The full proof of Lemma~\ref{lem:cut-reduction} is in the Appendix. Then we
have the following lemma.

@@

By induction on $c(\Pi)$ and Lemma~\ref{lem:less-cut-rank}, the cut
elimination theorem follows immediately.
\begin{theorem}[Cut Elimination]
  Let $\Pi$ be a proof of a sequent $[[I |-c X]]$ or $[[G |-l A]]$ s.t.
  $c(\Pi)>0$. Then there is an algorithm which yields a cut-free proof
  $\Pi'$ of the same sequent.
\end{theorem}
